\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-04-28 forwards-and-backwards.tex
peter edits to 7.1 and alison comments to 8

In this section we prove Theorem~\ref{thm:local_convergence},
which simultaneously scales the parameters $\theta$ and $N$ and the interaction 
kernel $\rho_F$ in the special case in which $r\equiv 1\equiv\gamma$, $q_{\theta}(x,dy)$
is isotropic with zero mean, the kernel
$\rho_F$ is Gaussian, and the scaling 
limit is a reaction-diffusion equation.

@@

Let us write $\mathcal{L}^\theta f(x) := \theta \int (f(y)-f(x))q_\theta(x,y) dy$ 
where $q_\theta$ is a Gaussian kernel of mean $0$ and variance $1/\theta$. 
We note that $\mathcal{L}^\theta$ is the generator of a continuous (time and 
space) random walk, which makes jumps of mean $0$ and variance $1/\theta$ at
rate $\theta$.

In what follows we write $\psi_t^{\epsilon, x}(y)$ for the solution of
\begin{equation}
    \partial_t \psi_t^{\epsilon,x}
    =
    \mathcal{L}^\theta \psi_t^{\epsilon, x},
    \label{AlmostHeatEquation}
\end{equation}
with initial condition 
$\psi_0^{\epsilon,x}(y) = \rho_\epsilon(y-x) =p_{\epsilon^2}(x,y)$.

To see why $\psi_t^{\epsilon,x}$ is useful,
first note that for any time-dependent function $\phi_t(x)$
with time derivative $\dot \phi_t(x) = \partial_t \phi_t(x)$,
\begin{multline}
\label{time dep mg prob}
    \langle \phi_t(x), \eta_t(dx) \rangle
    = 
    \langle \phi_0(x), \eta_0(dx) \rangle
    +
    M_t(\phi_\cdot(\cdot))
    +
    \int_0^t \big\langle \mathcal{L}^\theta \phi_s(x) 
        + \dot \phi_s(x), \eta_s(dx) \big\rangle ds
    \\  {}
    +
    \int_{0}^t \big\langle \phi_s(x) F(x, \eta_s) , \eta_s(dx) \big\rangle ds ,
\end{multline}
where $M_t(\phi_\cdot(\cdot))$ is a martingale (with respect to the natural filtration)
with quadratic variation given by~\eqref{eqn:prelimit_martingale_variation}
with $f$ replaced by $\phi_s(\cdot)$.
So, taking $\phi_s(\cdot)=\psi_{t-s}^{\epsilon,x}(\cdot)$ for $0 \le s \le t$,
\begin{eqnarray}
    \nonumber
    \rho_\epsilon*\eta_t(x)&=&\langle\psi^{\epsilon,x}_0(y),\eta_t(dy)\rangle
    \\
    &=&\langle\psi^{\epsilon,x}_t(y),\eta_0(dy)\rangle
    +\int_0^t\big\langle\psi^{\epsilon,x}_{t-s}(y)
    F\big(\rho_\epsilon*\eta_s(y)\big), \eta_s(dy)\big\rangle ds
    +M_t(x),
    \label{expn rhoepsilon} 
\end{eqnarray}
where $M_t(x)$ has mean zero and a second moment we can easily write down.

@@

We proceed by induction. 
Taking expectations in~(\ref{expn rhoepsilon}), using that $F$ is bounded above, and 
applying Gronwall's inequality to 
$\langle\psi^{\epsilon,x}_{t-s}, \eta_s\rangle$ 
we obtain~(\ref{moment bound rhoepsilon})
(without the second term on the right hand side) in the case $k=1$.
Moreover, rearranging~(\ref{expn rhoepsilon}) we find 
\begin{equation}
-\int_0^t\big\langle\psi^{\epsilon,x}_{t-s}(y)
F(\rho_\epsilon*\eta_s(y), \eta_s(dy)\big\rangle ds
=\langle\psi^{\epsilon,x}_t(y),\eta_0(dy)\rangle
-\langle \psi^{\epsilon,x}_0(y),\eta_t(dy)\rangle +M_t(x),
\end{equation} 
and taking expectations again, 
since $\langle\psi^{\epsilon,x}_0,\eta_t\rangle>0$, and $M_0(x)=0$,
this yields 
$$\IE\Big[-\int_0^t\langle\psi^{\epsilon,x}_{t-s}(y)F(\rho_\epsilon*\eta_s(y), 
\eta_s(dy)\rangle ds\Big|{\cal F}_0\Big]
\leq \widehat{C}\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle.$$
Since $F$ is bounded above, there exists a constant $K$ such that $|F|\leq K-F$ and so 
combined with the bound on $\IE[\langle\psi^{\epsilon, x}_0(y),\eta_t(dy)\rangle]$ just
obtained, this in turn yields 
$$\IE\Big[\int_0^t\big\langle\psi^{\epsilon,x}_{t-s}(y)
|F(\rho_\epsilon*\eta_s(y)|, 
\eta_s(dy)\big\rangle ds\Big|{\cal F}_0\Big]\leq 
\widetilde{C}\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle,$$
which is~(\ref{flandoli trick for rhoepsilon})
in the case $k=1$.

Now suppose that we have established~(\ref{moment bound rhoepsilon}) 
and~(\ref{flandoli trick for rhoepsilon}) for all exponents $j<k$.
First we apply the generator $\Pgen^N$ of our scaled population process
to functions of the form $\langle f,\eta\rangle^k$.
Recalling that each jump of the process involves the birth or death of a single 
individual, and so increments $\langle f,\eta\rangle$ by $\pm f/N$ at the location of
that individual, we find 
\begin{multline}
\label{pgen applied to kth moment}
\Pgen^N\Big(\langle f,\eta\rangle^k\Big)
=
\Big\langle
\int\theta N\Big\{k\frac{f(y)}{N}\langle f,\eta\rangle^{k-1}+
\binom{k}{2}\frac{f(y)^2}{N^2}\langle f,\eta\rangle^{k-2}+\cdots
+\frac{f(y)^{k}}{N^{k}}\Big\}q_\theta(x,dy),\eta(dx)\Big\rangle
\\
+
\Big\langle\theta N\Big(1-\frac{F(\rho_\epsilon*\eta(x))}{\theta}\Big)
\Big\{-k\frac{f(x)}{N}\langle f,\eta\rangle^{k-1}+
\binom{k}{2}\frac{f(x)^2}{N^2}\langle f,\eta\rangle^{k-2}+\cdots
+\frac{(-1)^{k}f(x)^{k}}{N^{k}}\Big\},\eta(dx)\Big\rangle.
\end{multline}
Mimicking what we did above, we set $f(\cdot)=\psi^{\epsilon,x}_t(\cdot)$ and 
write
\begin{multline}
\label{kth moment for time varying function}
\IE\Big[\langle \psi^{\epsilon,x}_0,\eta_t\rangle^k\Big|{\cal F}_0\Big]
=\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle^k
+\IE\Big[\int_0^t\Pgen^N\big(\langle\psi^{\epsilon,x}_{t-s}(y),\eta_s(dy)\rangle^k\big) ds
\\
-\int_0^t\langle k\dot{\psi}^{\epsilon,x}_{t-s}(y), \eta_s(dy)\rangle
\big\langle \psi^{\epsilon, x}_{t-s}(y),\eta_s(dy)\big\rangle^{k-1}ds
\Big|{\cal F}_0\Big].
\end{multline}
Using (from Lemma~\ref{PsiBoundHS}) 
that $\sup_s\|\psi^{\epsilon,x}_s(\cdot)\|_\infty=C/\epsilon^d$,
$N\epsilon^d>1$,
and our inductive
hypothesis, we find
\begin{multline*}
\IE\Big[
\Big\langle
\int_0^t\theta N\sum_{j=2}^{k}\binom{k}{j}\int
\frac{\psi_{t-s}^{\epsilon,x}(z)^j}{N^j}
\langle \psi^{\epsilon,x}_{t-s},\eta_s\rangle^{k-j}
q_\theta(y,dz),
\eta_s(dy)\Big\rangle ds 
\\
+
\int_0^t\theta N\sum_{j=2}^{k}\binom{k}{j}\frac{\psi_{t-s}^{\epsilon,x}(y)^j}{N^j}\langle \psi^{\epsilon,x}_{t-s},\eta_s\rangle^{k-j}
(-1)^j\Big(1-\frac{F(\rho_\epsilon*\eta_s(y))}{\theta} 
\Big),
\eta_s(dy)\Big\rangle ds\Big|{\cal F}_0\Big]
\\
\leq 
C\IE\Big[\Big\langle\int_0^t\sum_{j=2}^k\frac{\theta}{N\epsilon^d}
\Big(\frac{1}{(N\epsilon^d)^{j-2}}\Big)
	\Big\langle \psi^{\epsilon,x}_{t-s}(y)
\Big(2+\frac{|F(\rho_\epsilon*\eta_s(y))|}{\theta}\Big),\eta_s(dy)\Big\rangle
\langle\psi^{\epsilon,x}_{t-s},\eta_s\rangle^{k-j}ds\big|{\cal F}_0\Big]
\\
\leq
C'\frac{\theta}{N\epsilon^d}\sum_{j=1}^{k-1}
\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle^j
\leq
C''\frac{\theta}{N\epsilon^d}
\Big(\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle^k
+\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle\Big).
\end{multline*}
Combining this with~(\ref{pgen applied to kth moment})
and~(\ref{kth moment for time varying function}), using once again the fact that $F$
is bounded above, we find 
\begin{multline*}
\label{kth step of induction}
\IE\Big[\langle \psi^{\epsilon,x}_0,\eta_t\rangle^k\Big|{\cal F}_0\Big]
\leq
\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle^k
+\tilde{C}\IE\Big[\int_0^t
\langle\psi^{\epsilon, x}_{t-s}(y),\eta_s(dy)\rangle^k ds
\Big|{\cal F}_0\Big]
\\
+C''\frac{\theta}{N\epsilon^d}
\Big(\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle^k+
\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle\Big)
,
\end{multline*}
and~(\ref{moment bound rhoepsilon}) 
follows from Gronwall's inequality. Rearranging exactly as in the case $k=1$, we 
recover~(\ref{flandoli trick for rhoepsilon}) 
and the inductive step is complete.
\end{proof}

@@

To illustrate the inductive step, now consider
\begin{multline}
\label{rewrite for inductive step}
\IE\big[\langle \rho_\epsilon*\eta_t(x)^2,\eta_t(dx)\rangle\big] 
=
\IE\Big[\int\int\int p_{\epsilon^2}(x,z_1) p_{\epsilon^2}(x,z_2)
\eta_t(dz_1)\eta_t(dz_2)\eta_t(dx)\Big]
\\
=
\IE\Big[\int\cdots\int p_{\epsilon^2/2}(x,y_1) 
p_{\epsilon^2/2}(x,y_2)p_{\epsilon^2/2}(y_1,z_1)
p_{\epsilon^2/2}(y_2,z_2)
\eta_t(dz_1)\eta_t(dz_2)dy_1dy_2\eta_t(dx)\Big].
\end{multline}
We use the identity
\[
p_{\epsilon^2/2}(x,y_1) p_{\epsilon^2/2}(x,y_2)
=p_{\epsilon^2}(y_1,y_2)p_{\epsilon^2/4}\Big(x, \frac{y_1+y_2}{2}\Big)
\]
to rewrite~(\ref{rewrite for inductive step}) as
\begin{align*}
&\IE\Big[\int p_{\epsilon^2/2}*\eta_t(y_1)\, p_{\epsilon^2/2}*\eta_t(y_2)\,
p_{\epsilon^2/4}*\eta_y\big(\frac{y_1+y_2}{2}\big)\, 
p_{\epsilon^2}(y_1,y_2)dy_1dy_2\Big]
\\
&\leq 
\IE\Big[\int\Big\{
\big(p_{\epsilon^2/2}*\eta_t(y_1)\big)^3+ \big(p_{\epsilon^2/2}*\eta_t(y_2)\big)^3+
\big(p_{\epsilon^2/4}*\eta_t\big(\frac{y_1+y_2}{2}\big)\big)^3
\Big\}p_{\epsilon^2}(y_1,y_2)dy_1dy_2\Big],
\end{align*}
where we have used that for any non-negative real numbers
$\beta_1$, $\beta_2$, $\beta_3$, 
$\beta_1\beta_2\beta_3\leq \beta_1^3+\beta_2^3+\beta_3^3$.
For the first two terms in the sum we integrate with respect to $y_2$ and $y_1$
respectively to reduce to an expression of the form considered in 
Lemma~\ref{bounds on moments}. For the final term, the change of variables
$z_1=y_1+y_2$, $z_2=y_1-y_2$ in the integral similarly 
allows us to integrate out the heat kernel, and we conclude that the result
holds for $k=2$.

We can proceed in the same way for larger values of $k$, using repeatedly that
\[
p_{t_1}(x,y_1)p_{t_2}(x,y_2)=p_{\frac{t_1t_2}{t_1+t_2}}
\Big(x,\frac{t_2y_1+t_1y_2}{t_1+t_2}\Big)p_{t_1+t_2}(y_1,y_2)
\]
to write 
\[
\prod_{j=1}^k p_{\tau}(x,x_j)
=\prod_{j=2}^{k}p_{\frac{j\tau}{j-1}}\big(x_j,X_{j-1}\big)
p_{\frac{\tau}{k}}(x, X_k)
\]
where
\[
X_1=x_1, \qquad X_j=\frac{j-1}{j}x_j+\frac{1}{j}X_{j-1}, \mbox{for }j\geq 2.
\]
This yields
\begin{multline*}
\big\langle\big(\rho_\epsilon*\eta_t(x)\big)^k,\eta_t(dx)\big\rangle
=\int\cdots\int\prod_{i=2}^kp_{\epsilon^2 i/(i-1)}(y_,X_i)
\prod_{i=1}^kp_{\epsilon^2/2}*\eta_t(y_i)p_{\epsilon^2/k}*\eta_t(X_k)
dy_1\ldots dy_k
\\
\leq
\int\cdots\int 
\prod_{i=2}^kp_{\epsilon^2 i/(i-1)}(y_i,X_i)
\Big\{\sum_{i=1}^k\big(p_{\epsilon^2/2}*\eta_t(y_i)\big)^{k+1}
+\big(p_{\epsilon^2/k}*\eta_t(X_k)\big)^{k+1}\Big\}dy_1\ldots dy_k,
\end{multline*}
and once again we can change variables in the integrals and 
use~(\ref{heat equation at smaller time}) to bound this by a constant 
multiple of $\int\IE\big[\big(\rho_\epsilon*\eta_t(x)\big)^{k+1}\big]dx$, and
the inductive step is complete.
\end{proof}

@@

As in the nonlocal case, it suffices to prove that for $T>0$,
and any $f\in C_b^\infty (\IR^d)$ with bounded second derivatives, the
sequence of real-valued processes 
$\big\{\big(\int f(x)\rho_\epsilon*\eta_t^N(x)dx\big)_{t\geq 0}\big\}_{N\geq 1}$ is tight.  
Let us temporarily write $X_f^N(t)$ for $\int f(x)\rho_\epsilon*\eta_t^N(x)dx$
and set
\[
w'\big(X^N,\delta,T\big)= \inf_{\{t_i\}}\max_i\sup_{s,t\in [t_{i-1},t_i)}
\big| X_f^N(t)-X_f^N(s)\big|,
\]
where $\{t_i\}$ ranges over all partitions of the form 
$0=t_0<t_1<\cdots <t_{n-1}<T\leq t_n$ with 
$\min_{1\leq i\leq n}(t_i-t_{i-1})>\delta$ and $n\geq 1$.
Using Corollary~3.7.4 of~\cite{ethier/kurtz:1986}, 
to prove tightness
of the sequence of real-valued processes $X_f^N$
it suffices to check compact containment of 
the sequence $\{\rho_\epsilon*\eta_t^N(x)dx\}_{N\geq 1}$ at any rational time $t$ and
that for every $\nu>0$ and $T>0$, there exists $\delta>0$ such that
\[
\limsup_{N\to\infty}\IP\big[w'\big(X_f^N,\delta,T\big)>\nu\big]<\nu.
\]
Evidently this will follow 
if we can show that this condition is satisfied when we replace the 
minimum over all partitions with mesh at least $\delta$ in the 
definition of $w'$, by the partition into
intervals of length exactly $\delta$.

>>>>

In this section we prove Theorem~\ref{thm:local_convergence},
which proves convergence in the case that the width of the interaction kernel $\rho_F$
simultaneously scales along with the parameters $\theta$ and $N$,
in the special case in which $r\equiv 1\equiv\gamma$, $q_{\theta}(x,dy)$
is isotropic with zero mean, the kernel
$\rho_F$ is Gaussian, and the scaling 
limit is a reaction-diffusion equation.

@@

Let us write $\mathcal{L}^\theta f(x) := \theta \int (f(y)-f(x))q_\theta(x,y) dy$ 
where $q_\theta$ is a Gaussian kernel of mean $0$ and variance $1/\theta$. 
We note that $\mathcal{L}^\theta$ is the generator of a continuous (time and 
space) random walk, which makes jumps of mean $0$ and variance $1/\theta$ at rate $\theta$.
In what follows we write $\psi_t^{\epsilon, x}(y)$ for the solution of
\begin{equation}
    \partial_t \psi_t^{\epsilon,x}
    =
    \mathcal{L}^\theta \psi_t^{\epsilon, x},
    \label{AlmostHeatEquation}
\end{equation}
with initial condition 
$\psi_0^{\epsilon,x}(y) = \rho_\epsilon(y-x) =p_{\epsilon^2}(x,y)$.

To see why $\psi_t^{\epsilon,x}$ is useful,
first note that for any time-dependent function $\phi_t(x)$
with time derivative $\dot \phi_t(x) = \partial_t \phi_t(x)$,
\begin{multline}
\label{time dep mg prob}
    \langle \phi_t(x), \eta_t(dx) \rangle
    = 
    \langle \phi_0(x), \eta_0(dx) \rangle
    +
    M_t(\phi)
    +
    \int_0^t \big\langle \mathcal{L}^\theta \phi_s(x) 
        + \dot \phi_s(x), \eta_s(dx) \big\rangle ds
    \\  {}
    +
    \int_{0}^t \big\langle \phi_s(x) F(x, \eta_s) , \eta_s(dx) \big\rangle ds ,
\end{multline}
where $M_t(\phi)$ is a martingale (with respect to the natural filtration)
with quadratic variation given by~\eqref{eqn:prelimit_martingale_variation}
with $f$ replaced by $\phi_s(\cdot)$.
So, taking $\phi_s(\cdot)=\psi_{t-s}^{\epsilon,x}(\cdot)$ for $0 \le s \le t$,
\begin{eqnarray}
    \nonumber
    \rho_\epsilon*\eta_t(x)&=&\langle\psi^{\epsilon,x}_0(y),\eta_t(dy)\rangle
    \\
    &=&\langle\psi^{\epsilon,x}_t(y),\eta_0(dy)\rangle
    +\int_0^t\big\langle\psi^{\epsilon,x}_{t-s}(y)
    F\big(\rho_\epsilon*\eta_s(y)\big), \eta_s(dy)\big\rangle ds
    +M_t(x),
    \label{expn rhoepsilon} 
\end{eqnarray}
where $M_t(x)$ has mean zero and a second moment we can easily write down.

@@

We proceed by induction. 
Taking expectations in~(\ref{expn rhoepsilon}), using that $F$ is bounded above, and 
applying Gronwall's inequality to 
$\langle\psi^{\epsilon,x}_{t-s}, \eta_s\rangle$ 
we obtain 
$\IE[\langle \psi_0^{\epsilon,x},\eta_t\rangle] \le C \IE[\langle \psi_t^{\epsilon,x},\eta_0\rangle]$,
which implies~(\ref{moment bound rhoepsilon}) in the case $k=1$.
Moreover, rearranging~(\ref{expn rhoepsilon}) we find 
\begin{equation}
-\int_0^t\big\langle\psi^{\epsilon,x}_{t-s}(y)
F(\rho_\epsilon*\eta_s(y), \eta_s(dy)\big\rangle ds
=\langle\psi^{\epsilon,x}_t(y),\eta_0(dy)\rangle
-\langle \psi^{\epsilon,x}_0(y),\eta_t(dy)\rangle +M_t(x),
\end{equation} 
and taking expectations again, 
since $\langle\psi^{\epsilon,x}_0,\eta_t\rangle>0$, and $M_0(x)=0$,
this yields 
$$\IE\Big[-\int_0^t\langle\psi^{\epsilon,x}_{t-s}(y)F(\rho_\epsilon*\eta_s(y)), 
\eta_s(dy)\rangle ds\Big|{\cal F}_0\Big]
\leq \langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle.$$
Since $F$ is bounded above, there exists a constant $K$ such that $|F|\leq K-F$ and so 
combined with the bound on $\IE[\langle\psi^{\epsilon, x}_0(y),\eta_t(dy)\rangle]$ just
obtained, this in turn yields 
$$\IE\Big[\int_0^t\big\langle\psi^{\epsilon,x}_{t-s}(y)
|F(\rho_\epsilon*\eta_s(y))|, 
\eta_s(dy)\big\rangle ds\Big|{\cal F}_0\Big]\leq 
\widetilde{C}\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle,$$
which is~(\ref{flandoli trick for rhoepsilon})
in the case $k=1$.

Now suppose that we have established~(\ref{moment bound rhoepsilon}) 
and~(\ref{flandoli trick for rhoepsilon}) for all exponents $j<k$.
First we apply the generator $\Pgen^N$ of our scaled population process
to functions of the form $\langle f,\eta\rangle^k$.
Recalling that each jump of the process involves the birth or death of a single 
individual, and so increments $\langle f,\eta\rangle$ by $\pm f/N$ at the location of
that individual and that $r \equiv \gamma \equiv 1$, we find 
\begin{multline}
\label{pgen applied to kth moment}
\Pgen^N\Big(\langle f,\eta\rangle^k\Big)
\Big\langle
\int\theta N
    % \Big\{k\frac{f(y)}{N}\langle f,\eta\rangle^{k-1}+ \binom{k}{2}\frac{f(y)^2}{N^2}\langle f,\eta\rangle^{k-2}+\cdots +\frac{f(y)^{k}}{N^{k}}\Big\}
    \sum_{j=1}^k \binom{k}{j} \frac{f(y)^j}{N^j} \langle f, \eta\rangle^{k-j}
    q_\theta(x,dy),\eta(dx)\Big\rangle
    \Big\langle\theta N\Big(1-\frac{F(\rho_\epsilon*\eta(x))}{\theta}\Big)
    % \Big\{-k\frac{f(x)}{N}\langle f,\eta\rangle^{k-1}+ \binom{k}{2}\frac{f(x)^2}{N^2}\langle f,\eta\rangle^{k-2}+\cdots +\frac{(-1)^{k}f(x)^{k}}{N^{k}}\Big\}
    \sum_{j=1}^k \binom{k}{j} (-1)^j \frac{f(x)^j}{N^j} \langle f, \eta\rangle^{k-j}
    ,\eta(dx)\Big\rangle.
\end{multline}
Mimicking what we did above, we set $f(\cdot)=\psi^{\epsilon,x}_t(\cdot)$ and 
write
\begin{multline}
\label{kth moment for time varying function}
\IE\Big[\langle \psi^{\epsilon,x}_0,\eta_t\rangle^k\Big|{\cal F}_0\Big]
=\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle^k
+\IE\Big[\int_0^t\Pgen^N\big(\langle\psi^{\epsilon,x}_{t-s}(y),\eta_s(dy)\rangle^k\big) ds
-\int_0^t\langle k\dot{\psi}^{\epsilon,x}_{t-s}(y), \eta_s(dy)\rangle
\big\langle \psi^{\epsilon, x}_{t-s}(y),\eta_s(dy)\big\rangle^{k-1}ds
\Big|{\cal F}_0\Big].
\end{multline}
Since $\dot \psi^{\epsilon,x}_s = \mathcal{L}_\theta \psi^{\epsilon,x}_s$,
the $j=1$ terms from $\Pgen^N(\langle \psi^{\epsilon,x}_{t-s}(y),\eta_s(dy) \rangle^k)$
combines with the last term in~\eqref{kth moment for time varying function}
to obtain 
$$ \int_0^t k \langle \psi^{\epsilon,x}_{t-s},\eta\rangle^{k-1} \langle F(\rho_\epsilon*\eta_s(y)) \psi^{\epsilon,x}_{t-s}(y), \eta_s(dy) \rangle ds . $$
As for the remaining terms, using (from Lemma~\ref{PsiBoundHS}) 
that $\sup_s\|\psi^{\epsilon,x}_s(\cdot)\|_\infty=C/\epsilon^d$,
$N\epsilon^d>1$,
and our inductive
hypothesis, we find
\begin{multline*}
\IE\Big[
\Big\langle
\int_0^t\theta N\sum_{j=2}^{k}\binom{k}{j}\int
\frac{\psi_{t-s}^{\epsilon,x}(z)^j}{N^j}
\langle \psi^{\epsilon,x}_{t-s},\eta_s\rangle^{k-j}
q_\theta(y,dz),
\eta_s(dy)\Big\rangle ds 
\Big\langle \int_0^t\theta N\sum_{j=2}^{k}\binom{k}{j}\frac{\psi_{t-s}^{\epsilon,x}(y)^j}{N^j}\langle \psi^{\epsilon,x}_{t-s},\eta_s\rangle^{k-j}
(-1)^j\Big(1-\frac{F(\rho_\epsilon*\eta_s(y))}{\theta} 
\Big),
\eta_s(dy)\Big\rangle ds\Big|{\cal F}_0\Big]
C\IE\Big[\Big\langle\int_0^t\sum_{j=2}^k\frac{\theta}{N\epsilon^d}
\Big(\frac{1}{(N\epsilon^d)^{j-2}}\Big)
	\Big\langle \psi^{\epsilon,x}_{t-s}(y)
\Big(2+\frac{|F(\rho_\epsilon*\eta_s(y))|}{\theta}\Big),\eta_s(dy)\Big\rangle
\langle\psi^{\epsilon,x}_{t-s},\eta_s\rangle^{k-j}ds\big|{\cal F}_0\Big]
C'\frac{\theta}{N\epsilon^d}\sum_{j=1}^{k-1}
\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle^j
C''\frac{\theta}{N\epsilon^d}
\Big(\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle^k
+\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle\Big).
\end{multline*}
Combining this with~(\ref{pgen applied to kth moment})
and~(\ref{kth moment for time varying function}), using once again the fact that $F$
is bounded above, we find 
\begin{multline*}
\label{kth step of induction}
\IE\Big[\langle \psi^{\epsilon,x}_0,\eta_t\rangle^k\Big|{\cal F}_0\Big]
\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle^k
+\tilde{C}\IE\Big[\int_0^t
\langle\psi^{\epsilon, x}_{t-s}(y),\eta_s(dy)\rangle^k ds
\Big|{\cal F}_0\Big]
+C''\frac{\theta}{N\epsilon^d}
\Big(\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle^k+
\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle\Big)
,
\end{multline*}
and~(\ref{moment bound rhoepsilon}) 
follows from Gronwall's inequality. Rearranging exactly as in the case $k=1$, we 
recover~(\ref{flandoli trick for rhoepsilon}) 
and the inductive step is complete.

@@

To illustrate the inductive step, now consider
\begin{multline}
\label{rewrite for inductive step}
\IE\big[\langle \rho_\epsilon*\eta_t(x)^2,\eta_t(dx)\rangle\big] 
=
\IE\Big[\int\int\int p_{\epsilon^2}(x,z_1) p_{\epsilon^2}(x,z_2)
\eta_t(dz_1)\eta_t(dz_2)\eta_t(dx)\Big]
\\
=
\IE\Big[\int\cdots\int p_{\epsilon^2/2}(x,y_1) 
p_{\epsilon^2/2}(x,y_2)p_{\epsilon^2/2}(y_1,z_1)
p_{\epsilon^2/2}(y_2,z_2)
\eta_t(dz_1)\eta_t(dz_2)dy_1dy_2\eta_t(dx)\Big].
\end{multline}
We use the identity
\[
p_{\epsilon^2/2}(x,y_1) p_{\epsilon^2/2}(x,y_2)
=p_{\epsilon^2}(y_1,y_2)p_{\epsilon^2/4}\Big(x, \frac{y_1+y_2}{2}\Big)
\]
to rewrite~(\ref{rewrite for inductive step}) as
\begin{align*}
&\IE\Big[\int \int p_{\epsilon^2/2}*\eta_t(y_1)\, p_{\epsilon^2/2}*\eta_t(y_2)\,
p_{\epsilon^2/4}*\eta_t\big(\frac{y_1+y_2}{2}\big)\, 
p_{\epsilon^2}(y_1,y_2)dy_1dy_2\Big]
\\
&\leq 
\IE\Big[\int \int\Big\{
\big(p_{\epsilon^2/2}*\eta_t(y_1)\big)^3+ \big(p_{\epsilon^2/2}*\eta_t(y_2)\big)^3+
\big(p_{\epsilon^2/4}*\eta_t\big(\frac{y_1+y_2}{2}\big)\big)^3
\Big\}p_{\epsilon^2}(y_1,y_2)dy_1dy_2\Big],
\end{align*}
where we have used that for any non-negative real numbers
$\beta_1$, $\beta_2$, $\beta_3$, 
$\beta_1\beta_2\beta_3\leq \beta_1^3+\beta_2^3+\beta_3^3$.
For the first two terms in the sum we integrate with respect to $y_2$ and $y_1$
respectively to reduce to an expression of the form considered in 
Lemma~\ref{bounds on moments}. For the final term, the change of variables
$z_1=y_1+y_2$, $z_2=y_1-y_2$ in the integral similarly 
allows us to integrate out the heat kernel, and we conclude that the result
holds for $k=2$.

We can proceed in the same way for larger values of $k$, using repeatedly that
\[
p_{t_1}(x,y_1)p_{t_2}(x,y_2)=p_{\frac{t_1t_2}{t_1+t_2}}
\Big(x,\frac{t_2y_1+t_1y_2}{t_1+t_2}\Big)p_{t_1+t_2}(y_1,y_2)
\]
to write 
\[
\prod_{j=1}^k p_{\tau}(y,y_j)
=\prod_{j=2}^{k}p_{\frac{j\tau}{j-1}}\big(y_j,Y_{j-1}\big)
p_{\frac{\tau}{k}}(y, Y_k)
\]
where
\[
    Y_1=y_1, \qquad Y_j=\frac{j-1}{j}Y_{j-1}+\frac{1}{j}y_j,  \mbox{for }j\geq 2.
\]
Writing $p_{\epsilon^2}(x,z_j) = \int p_{\epsilon^2/2}(x,y_j) p_{\epsilon^2/2}(y_j,z_j) dy_j$
and using the above with $\tau = \epsilon^2/2$,
this yields
\begin{multline*}
\big\langle\big(\rho_\epsilon*\eta_t(x)\big)^k,\eta_t(dx)\big\rangle
    =\int\cdots\int \prod_{j=2}^k p_{\epsilon^2 j/2(j-1)}(y_j,Y_{j-1})
     \prod_{i=1}^k p_{\epsilon^2/2}*\eta_t(y_i)p_{\epsilon^2/2k}*\eta_t(Y_k)
dy_1\ldots dy_k
\\
\leq
\int\cdots\int 
    \prod_{j=2}^k p_{\epsilon^2 j/2(j-1)}(y_j,Y_{j-1})
    \Big\{\sum_{i=1}^k\big(p_{\epsilon^2/2}*\eta_t(y_i)\big)^{k+1}
    + \big(p_{\epsilon^2/2k}*\eta_t(Y_k)\big)^{k+1}\Big\}dy_1\ldots dy_k,
\end{multline*}
and once again we can change variables in the integrals and 
use~(\ref{heat equation at smaller time}) to bound this by a constant 
multiple of $\int\IE\big[\big(\rho_\epsilon*\eta_t(x)\big)^{k+1}\big]dx$, and
the inductive step is complete.
\end{proof}

@@

As in the nonlocal case, it suffices to prove that for $T>0$,
and any $f\in C_b^\infty (\IR^d)$ with bounded second derivatives
and $\int|f(x)|dx < \infty$, the sequence of real-valued processes 
$\big\{\big(\int f(x)\rho_\epsilon*\eta_t^N(x)dx\big)_{t\geq 0}\big\}_{N\geq 1}$ is tight.  
Let us temporarily write $X_f^N(t)$ for $\int f(x)\rho_\epsilon*\eta_t^N(x)dx$
and set
\[
w'\big(X_f^N,\delta,T\big)= \inf_{\{t_i\}}\max_i\sup_{s,t\in [t_{i-1},t_i)}
\big| X_f^N(t)-X_f^N(s)\big|,
\]
where $\{t_i\}$ ranges over all partitions of the form 
$0=t_0<t_1<\cdots <t_{n-1}<T\leq t_n$ with 
$\min_{1\leq i\leq n}(t_i-t_{i-1})>\delta$ and $n\geq 1$.
Using Corollary~3.7.4 of~\cite{ethier/kurtz:1986}, 
to prove tightness
of the sequence of real-valued processes $X_f^N$
it suffices to check compact containment of 
    the sequence $\{\int f(x) \rho_\epsilon*\eta_t^N(x)dx\}_{N\geq 1}$ at any rational time $t$ and
that for every $\nu>0$ and $T>0$, there exists $\delta>0$ such that
\[
\limsup_{N\to\infty}\IP\big[w'\big(X_f^N,\delta,T\big)>\nu\big]<\nu.
\]
Evidently this will follow 
if we can show that this condition is satisfied when we replace the 
minimum over all partitions with mesh at least $\delta$ in the 
definition of $w'$, by the partition into
intervals of length exactly $\delta$.

We have
\begin{multline}
\label{tightness estimate}
\left|
\langle \rho_{\epsilon} * f, \eta_t^N \rangle 
- \langle \rho_{\epsilon} * f, \eta_s^N \rangle            
\right|
\leq
\left|
\int_s^t\Big\langle\theta\int\big(\rho_\epsilon*f(y)-\rho_\epsilon*f(x)\big)
q_{\theta}(x,dy),\eta_u^N(dx)\Big\rangle du
\right|
\\
+\int_s^t\Big\langle |F\big(\rho_\epsilon*\eta_u^N(x)\big)|\rho_\epsilon*|f|(x), 
\eta_u^N(dx)\Big\rangle du +2\sup_{0\leq u\leq T}|\widehat{M}^N(f)_u|,
\end{multline}
where $\widehat{M}^N(f)$ is the martingale of~(\ref{eqn:eta_f_mgale_decomp})
with the test function $f$ replaced by $\rho_\epsilon*f$. 
We control each of the three terms on the right hand side separately.
